\documentclass{article}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white

\begin{document}

\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}

\lstset{language=Matlab,%
    basicstyle=\footnotesize\ttfamily,
    breaklines=false,%
    morekeywords={matlab2tikz},
    keywordstyle=\color{blue},%
    morekeywords=[2]{1}, keywordstyle=[2]{\color{black}},
    identifierstyle=\color{black},%
    stringstyle=\color{mylilas},
    commentstyle=\color{mygreen},%
    showstringspaces=false,%without this there will be a symbol in the places where there is a space
    emph=[1]{for,end,break},emphstyle=[1]\color{red}, %some words to emphasise
    %emph=[2]{word1,word2}, emphstyle=[2]{style},
}

\begin{centering}
	{\scshape\Large FMA240 - Handin 3\par}
	\vspace{0.5cm}
	Kristoffer Lundgren \texttt{<kem01klu@student.lu.se>}\par
	Stefan Eng \texttt{<atn08sen@student.lu.se>}\par
    \vspace{0.5cm}
	\today\par
    \rule{\textwidth}{0.4pt}
\end{centering}

\section*{Exercise 1}
  In a transportation problem we can always assume that the total supply equals
  the total demand. If this is not the case, one can always add a dummy column
  or row with 0 costs to obtain balance by supplying the missing demand or
  supply. The general transportation problem can then be formulated as
  \begin{align*}
  & \text{Minimize } z = \sum_{i=1}^{m}\sum_{j=1}^n c_{ij}x_{ij} \\
  & \text{subject to}
    \label{eq1}
  \end{align*}
  \begin{align}
    \begin{cases}
      \sum_{j=1}^{n} x_{ij} = s_{i}, & i = 1, 2, ..., m \\
      \sum_{i=1}^{m} x_{ij} = d_{j}, & j = 1, 2, ..., n
    \end{cases}
    %\label{eq2}
  \end{align}
  \begin{equation}
    \sum_{i=1}^{m} s_{i} = \sum_{j=1}^n d_{j}
    \label{eq3}
  \end{equation}
  \begin{equation}
    x_{ij} \geq 0, i = 1, 2, ..., m, j = 1, 2, ..., n.
    \label{eq4}
  \end{equation}

  \noindent
  In the case of this handin, the cost matrix is given by
  \begin{equation}
    \mathbf{C} =
    \begin{pmatrix}
      11 & * & 8 & 8 \\
      7  & 5 & 6 & 12 \\
      7  & 6 & 8 & 5
    \end{pmatrix},
    \label{eq5}
  \end{equation}
  where $*$ signifies that supplier 1 cannot deliver to location 2. Also,
  supplier 1 offers a discount for location 4: every unit after $20k$ costs
  only 5, instead of 8.
  The supply vector is give by
  \begin{equation}
    \mathbf{s} =
    \begin{pmatrix}
      100k & 120k & 60k
    \end{pmatrix}^T
    \label{eq6}
  \end{equation}
  where $k = 10^3$. Finally, the demand vector is given by
  \begin{equation}
    \mathbf{d} =
    \begin{pmatrix}
      50k & 40k & 90k & 70k
    \end{pmatrix}^T,
    \label{eq7}
  \end{equation}
  where again $k = 10^3$. Obviously, for this case it does not hold that supply
  equals demand. In fact, there is $30k$ less demand than there is supply. As
  mentioned above, this can be taken care of by adding a dummy delivery
  location, which soaks up the remaining $30k$, with
  a unit transportation cost of $0$. This means that the cost matrix in
  equation (\ref{eq5}) becomes
  \begin{equation}
    \mathbf{C} =
    \begin{pmatrix}
      11 & * & 8 & 8  & 0\\
      7  & 5 & 6 & 12 & 0\\
      7  & 6 & 8 & 5  & 0
    \end{pmatrix},
    \label{eq8}
  \end{equation}
  as well as the demand vector in equation (\ref{eq7}) changing to
  \begin{equation}
    \mathbf{d} =
    \begin{pmatrix}
      50k & 40k & 90k & 70k & 30k
    \end{pmatrix}^T,
    \label{eq9}
  \end{equation}.

  \noindent
  To make sure that nothing gets delivered to location 2 from supplier 1, the
  constraints in equation (\ref{eq4}) must be modified such that for $x_{12}$
  it must be that $x_{12} = 0$. The discount in turn can handled by dividing
  the problem in two cases: either location 4 buys 0 to 20k units from supplier
  1 or more than 20k units from supplier 1. To mathematically formulate
  optimization problems for these two cases, location 4 is divided into two
  locations, 4a and 4b. In case 1, location 4b cannot buy from supplier 1, in
  case 2 location 4a cannot buy from supplier 2 and 3. The best solution for
  either of these two cases is then accepted as the answer to the original
  problem. To formulate this more mathematically, we thus have, for case 1:
 % m = 3
 % n = 6
  \begin{align*}
  & \text{Minimize } z = \sum_{i=1}^{3}\sum_{j=1}^6 c_{ij}x_{ij} \\
  & \text{subject to}
    \label{eq10}
  \end{align*}
  \begin{align}
    \begin{cases}
      \sum_{j=1}^{6} x_{ij} = s_{i}, & i = 1, 2, 3 \\
      \sum_{i=1}^{3} x_{ij} = d_{j}, & j = 1, 2, ..., 6
    \end{cases}
    %\label{eq11}
  \end{align}
  \begin{equation}
    \sum_{i=1}^{3} s_{i} = \sum_{j=1}^6 d_{j}
    \label{eq12}
  \end{equation}
  \begin{equation}
    x_{ij} \geq 0, i = 1, 2, 3, j = 1, 2, ..., 6, x_{12} = 0 x_{15} = 0.
    \label{eq13}
  \end{equation}
  with
  \begin{equation}
    \mathbf{C} =
    \begin{pmatrix}
      11 & * & 8 & 8  & *  & 0\\
      7  & 5 & 6 & 12 & 12 & 0\\
      7  & 6 & 8 & 5  & 5  & 0
    \end{pmatrix},
    \label{eq14}
  \end{equation}
  \begin{equation}
    \mathbf{d} =
    \begin{pmatrix}
      50k & 40k & 90k & 20k & 50k & 70k & 30k
    \end{pmatrix}^T
    \label{eq15}
  \end{equation}
  and
  \begin{equation}
    \mathbf{s} =
    \begin{pmatrix}
      100k & 120k & 60k
    \end{pmatrix}^T.
    \label{eq16}
  \end{equation}

  \noindent
  For case 2 we instead have:
   \begin{align*}
  & \text{Minimize } z = \sum_{i=1}^{3}\sum_{j=1}^6 c_{ij}x_{ij} \\
  & \text{subject to}
    \label{eq17}
  \end{align*}
  \begin{align}
    \begin{cases}
      \sum_{j=1}^{6} x_{ij} = s_{i}, & i = 1, 2, 3 \\
      \sum_{i=1}^{3} x_{ij} = d_{j}, & j = 1, 2, ..., 6
    \end{cases}
    %\label{eq18}
  \end{align}
  \begin{equation}
    \sum_{i=1}^{3} s_{i} = \sum_{j=1}^6 d_{j}
    \label{eq19}
  \end{equation}
  \begin{equation}
    x_{ij} \geq 0, i = 1, 2, 3, j = 1, 2, ..., 6, x_{12} = 0, x_{24} = 0, x_{34} = 0.
    \label{eq20}
  \end{equation}
  with
  \begin{equation}
    \mathbf{C} =
    \begin{pmatrix}
      11 & * & 8 & 8  & 5  & 0\\
      7  & 5 & 6 & *  & 12 & 0\\
      7  & 6 & 8 & *  & 5  & 0
    \end{pmatrix},
    \label{eq21}
  \end{equation}
  \begin{equation}
    \mathbf{d} =
    \begin{pmatrix}
      50k & 40k & 90k & 20k & 50k & 70k & 30k
    \end{pmatrix}^T
    \label{eq22}
  \end{equation}
  and
  \begin{equation}
    \mathbf{s} =
    \begin{pmatrix}
      100k & 120k & 60k
    \end{pmatrix}^T.
    \label{eq23}
  \end{equation}

  \section*{Exercise 2}
  To solve the transportation problems formulated in exercise 1 above, it is
  easiest to use the transportation algorithm (although, writing out all the
  constraints and running for example simplex would be possible, but more work).

  \noindent
  The transportation algorithm starts by finding a feasible solution to the
  problem by some method. One is for example the \textit{Minimum cell cost}
  method, which is a greedy algorithm occupying the transport routes that are
  the cheapest for each row in the cost matrix and meeting the respective
  supply and demand, starting with the first row and then doing the same for
  each consecutive row. The rest of the routes are set to 0.
  To find an improvement to this starting solution (or verifying that it is the
  optimal solution), one then defines the reduced costs as
  \begin{equation}
    \hat{c}_{ij} = c_{ij} - v_i - w_j,
    \label{eq24}
  \end{equation}
  where $\hat{c}_{ij}$ would appear as slack variables when solving the dual
  system, $v_i$ and $w_j$ in turn are the variables for the dual system.
  Complementary slackness can now be used, which states that for an optimal
  solution it must hold that
  \begin{equation}
    \hat{c}_{ij}x_{ij} = 0 \Rightarrow \text{if } x_{ij} \neq 0 \text{ then } \hat{c}_{ij} = 0.
  \end{equation}
  Creating a table for the $\hat{c}_{ij}$:s, these must be 0 where $x_{ij}$ is
  not. For these positions, it must then hold that $c_{ij} = v_i + w_j$ (as can
  be seen from equation (\ref{eq24})). Solving this equation system, the rest
  of the reduced costs can be calculated. The reduced costs should then be
  interpreted as the change in the total cost that we are optimizing for if one
  more unit is transported along that route; thus, finding a negative reduced
  cost means that the solution to the transport problem can be improved.
  Repeating this procedure until no more negative reduced costs are found, the
  optimal solution to the problem has then been found and can be returned as
  the answer.

  \section*{Exercise 3}
  Solving the two problems listed under the section ``Exercise 1'' above, the
  respective solutions are:

  \subsection*{Case 1}
  \begin{lstlisting}
>> ca = [11    100     8   8       100     0;
         7     5       6   12      12      0;
         7     6       8   5       5       0];
>> da = [50;   40;     90; 20;     50;     30];
>> sa = [100; 120; 60];
>> [xa, costa] = transport(sa, da, ca)
Optimal flow found:

xa =

     0     0    60    10     0    30
    50    40    30     0     0     0
     0     0     0    10    50     0


costa =

        1590
  \end{lstlisting}

  \subsection*{Case 2}
  \begin{lstlisting}
>> cb = [11    100     8   8       5       0;
         7     5       6   100     12      0;
         7     6       8   100     5       0];
>> db = [50;   40;     90; 20;     50;     30];
>> sb = [100; 120; 60];
>> [xb, costb] = transport(sb, db, cb)
Optimal flow found:

xb =

     0     0     0    20    50    30
     0    30    90     0     0     0
    50    10     0     0     0     0


costb =
        1510
  \end{lstlisting}

  \noindent
  The total cost is thus lower in the case where location 4 buys more than
  $20k$ units from supplier 1 and thereby obtaining the discount. In summary,
  location 1 should by $50k$ units from supplier 3, location 2 should by $30k$
  units from supplier 2 and $10k$ units from supplier 3, location 3 should by
  $90k$ units from supplier 2 and location 4 should by $70k$ units from
  supplier 1, in order to minimize the total costs. The total cost is 1510000.

\end{document}
